---
title: "Final Project: The Effect of Emotion on Reactions to Greta Thunberg’s Tweets"
author: "Lily Franks, Sophia Giustino, Emma Kim, and Rob Dreyer"
date: '2022-12-18'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(tidytext)
library(Matrix)
library(rARPACK)
library(glmnet)
library(randomForest)
library(textdata)
library(rtweet)
library(dplyr)
```

Check out and download our full project on [Github](https://github.com/sgiustino/Stat433FinalProject).

## Introduction

Greta Thunberg is a Swedish environmental activist who is known for her work that addresses the issue of climate change. She has challenged world leaders to take immediate action against climate change and has even spoken at the United Nations Climate Action Summit in 2019. Greta joined the spotlight at the age of 18, and her Twitter account shortly followed. Today, she has over 4.9 million followers and more than 11.7 thousand tweets that discuss climate change and environmental activism. 

With the impending threats of climate change over the past half-decade, and the prevalence of social media today, we thought it would be interesting to investigate how social media synchronizes with current events; more specifically, how the use of social media could affect responses to climate change. We decided to do an in-depth analysis of Greta’s tweets to see if we could find any connections between the word choice in her tweets to the interactions with the tweet. We determined how the sentiment of social media posts about climate change influences engagement, which could lead to more actions being taken to stop climate change. We have found that Greta Thunberg’s tweets that have a higher ratio of emotionally-laden words to total words will receive greater interaction, measured by the number of retweets. 

## Data

To collect data from Greta Thunberg’s twitter account, we used pulled data from the Twitter API with the package `rtweet`. Using the `get_timeline` function, we were able to pull approximately 3000 of Greta Thunberg’s most recent tweets and write them to a csv file. We then filtered the Twitter API data to include the full text, retweet count, favorited count, language, and word count. To determine the sentiment in each tweet, we used three word dictionaries known as Bing, Loughran, and NRC. Each dictionary classifies a word as either positive, negative, or neutral. Some dictionaries had other classifications, but for consistency purposes we only used the positive and negative variables.
	
Before we analyzed the sentiment of each tweet, we first filtered the data to only include tweets written in English. We then split each tweet to include one word per row using the `unnest_tokens` function. With the tweets separated by word, we created three new data frames by joining each of the sentiment dictionaries onto the individual words of each tweet. Every word that was matched to a sentiment dictionary added to the count of positive or negative words in the tweet. By taking the sum of positive and negative words (i.e. `bingPos`, `bingNeg`) in each tweet and dividing by the word count, we found the ratio of emotionally-laden words in each tweet (i.e. `bingRatio`). We then used a left join to combine the three sentiment data frames onto the original dataset, allowing us to compare the sentiment data to the twitter interactions data. To make sure all rows were in order, we used the function, `mean(xsent$tweet == 1:nrow(xsent))`, to double check our left joins. 


```{r, echo = F}
## This is how we pulled the tweets from Twitter, but we only did this once since it automatically pulls the most recent tweets and we wanted to get consistent results.

#library(rtweet)
#tweets = get_timeline(user = "GretaThunberg", n = Inf)
#write.table(tweets, file = "/Users/sophiagiustino/Documents/Fall2022/Stat433/Project/greta_tweets.csv")
#tweets_cut = tweets %>% 
    #select(created_at, id, full_text, truncated, id_str, display_text_range,        #is_quote_status, retweet_count, favorite_count, favorited, retweeted, lang,     #text) 
#write_csv(tweets_cut, file = "/Users/sophiagiustino/Documents/Fall2022/Stat433/Project/greta_tweets.csv")
```


```{r import data, echo = F, include=F}
greta_tweets = read.csv("greta_tweets.csv")

text_df <- tibble(tweet = 1:nrow(greta_tweets), text = greta_tweets$full_text)

extra_text_df <- tibble(tweet = 1:nrow(greta_tweets), text = greta_tweets$full_text, retweets = greta_tweets$retweet_count, favorites = greta_tweets$favorite_count)

tweet1 = greta_tweets$full_text[1]
```

```{r data processing, echo = F, include=F}
# this does a lot of processing! 

tt  = text_df %>% 
  unnest_tokens(word, text) %>% 
  group_by(tweet) 

tt2  = text_df %>% 
  unnest_tokens(word, text) %>% 
  group_by(tweet) %>%
  summarise(word_count = n())

```


```{r bing sentiment, echo = F, include=F}
bing = get_sentiments("bing") %>% 
  mutate(neg= sentiment == "negative", 
         pos = sentiment =="positive")

bingFeat = tt %>% 
  left_join(bing) %>% 
  group_by(tweet) %>%
  summarize(bingNeg = sum(neg, na.rm = T),
            bingPos = sum(pos, na.rm = T),
            bingWords = bingNeg + bingPos,
            #bingRatio = bingWords/(tt$count),
            bing = (bingPos + .00000001)/(n()+.00000002))
```

```{r lough sentiment, echo = F, include=F}
lough = get_sentiments("loughran") %>% 
  rename(lough = sentiment)

lough = lough %>% 
  mutate(value = 1) %>% 
  pivot_wider(word, names_from = lough)


loughFeat = tt %>% 
  left_join(lough) %>% 
  group_by(tweet) %>% 
  summarize(  
    loughNeg = sum(negative, na.rm=T),
    loughPos = sum(positive, na.rm=T),
    loughWords = loughNeg + loughPos,
    #loughRatio = loughWords / count,
    lough = (loughPos + .00000001)/(n()+.00000002))
```

```{r nrc sentiment, echo = F, include=F}
nrc = get_sentiments("nrc") %>% 
  rename(nrc = sentiment)

nrcWide = nrc %>% 
  mutate(value = 1) %>% 
  pivot_wider(word, names_from = nrc, values_fill = 0)

nrcFeat = tt %>% 
  left_join(nrcWide) %>% 
  group_by(tweet) %>% 
  summarize(
    nrcNeg = sum(negative, na.rm=T),
    nrcPos = sum(positive, na.rm=T),
    nrcWords = nrcNeg + nrcPos,
    #nrcRatio = nrcWords / count,
    nrc = (nrcPos + .00000001)/(n()+.00000002)) 
```


```{r joining, echo = F, include=F}
xsent = tt2 %>%
  group_by(tweet) %>%
  left_join(bingFeat) %>% 
  left_join(loughFeat) %>% 
  left_join(nrcFeat) 
xsent[is.na(xsent)] = 0
```


```{r order check, echo = F, include=F}
#make sure rows are in order: 
mean(xsent$tweet == 1:nrow(xsent))
```

________________________


```{r create gret_sent, echo = F, include=F}
# THIS JOINS GRETATWEETS AND XSENT = GRET_SENT
gret_sent <- cbind(tweet = 1:nrow(greta_tweets), greta_tweets)
```

```{r clean up gret_sent, echo = F, include=F}
gret_sent = gret_sent %>%
  select(tweet, full_text, retweet_count, favorite_count, lang) %>%
  left_join(xsent) %>%
  mutate(bingRatio = bingWords/word_count, loughRatio = loughWords/word_count, nrcRatio = nrcWords/word_count) %>%
  filter(lang == "en") #just english!

gret_sent[is.na(gret_sent)] = 0
#write.csv(gret_sent, "/Users/sophiagiustino/Documents/Fall2022/Stat433/Project/gret_sent.csv", row.names = T)
```

## Methods

Upon completion of the dataframe `gret_sent`, we decided to use the ratio variables (i.e. `bingRatio`) as the independent variables in our study to analyze emotionally-laden words in correlation with retweet behavior (i.e. `retweet_count`). We did this because we wanted to measure emotional saturation of a tweet rather than the number of emotionally-laden words (i.e `bingWords`) within the tweet in order to support the ranging lengths of the tweets. Additionally, we chose to focus on all the emotional words instead of either positive or negative because all dictionary words carry emotional significance, which answers the question of our study. 

Next, in order to ensure all dictionaries (Bing, Lough, and NRC) were significantly correlated to the amount of retweets per tweet, we constructed three separate linear regression models along with anova tables. These tables proved to us that all three dictionary ratios had a statistically significant effect of retweet count. However, in order for these tables to portray statistically significant p values for the ratios we used the log value of `retweet_count`. In doing so, the graphs produced became more concise and the linear lines more dramatic. 

Lastly, we used the following equation to interpret the three dictionaries’ coefficients, within the anova tables: ln(Yi) = 𝜷0 + 𝜷1Xi + ui. While the equation originally interprets the relationship as a correlation between a tweet with zero emotionally-laden words in comparison to one with 100% emotionally-laden words, we decided to interpret these results with 10% increases. We did this in hopes to clarify the results in a real-world sense as there are rarely tweets that are completely made of emotional words.


## Results

We plotted the linear relationships between the emotional saturation of tweets and the logarithm of their retweet count. All of the graphs show a slight positive slope for the linear trend. You can also view these plots in our [Shiny App.](https://github.com/sgiustino/Stat433FinalProject/blob/master/app.R)

#### Figure 1: NRC Ratio vs. Retweet Count
```{r nrc graph, echo = F}
ggplot(gret_sent, aes(nrcRatio, log(retweet_count))) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm") +
  xlab("Ratio Measure (NRC Words: Total Tweet Words)") +
  ylab("Retweet Count (Log)") +
  ggtitle("How NRC Dict. Words Affect Retweet Count")
```

#### Table 1
```{r nrc reg, echo = F}
summary(lm(log(retweet_count) ~ nrcRatio, gret_sent))
```
When we regressed the logarithm of retweets on the ratio of NRC dictionary emotional words to total words in the tweet, we found the coefficient for the ratio of emotional words was statistically significant, with a coefficient of 1.19. This means that for a 10% increase in emotional words (an increase of 0.1 in the ratio), there is an 11.9% increase in retweets. 


#### Figure 2: Lough Ratio vs. Retweet Count
```{r lough graph, echo = F}
ggplot(gret_sent, aes(loughRatio, log(retweet_count))) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm") +
  xlab("Ratio Measure (Lough Words: Total Tweet Words)") +
  ylab("Retweet Count (Log)") +
  ggtitle("How Lough Dict. Words Affect Retweet Count")
```

#### Table 2
```{r lough reg, echo = F}
summary(lm(log(retweet_count) ~ loughRatio, gret_sent))
```
When we regressed the logarithm of retweets on the ratio of lough dictionary emotional words to total words in the tweet, we found the coefficient for the ratio of emotional words was also statistically significant, with a coefficient of 3.39. This means that for a 10% increase in emotional words (an increase of 0.1 in the ratio), there is an 33.9% increase in retweets.


#### Figure 3: Bing Ratio vs. Retweet Count
```{r bing graph, echo = F}
ggplot(gret_sent, aes(bingRatio, log(retweet_count))) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm") +
  xlab("Ratio Measure (Bing Words: Total Tweet Words)") +
  ylab("Retweet Count (Log)") +
  ggtitle("How Bing Dict. Words Affect Retweet Count")
```

#### Table 3
```{r bing reg, echo = F}
summary(lm(log(retweet_count) ~ bingRatio, gret_sent))
```

When we regressed the logarithm of retweets on the ratio of bing dictionary emotional words to total words in the tweet, we found the coefficient for the ratio of emotional words was also statistically significant, with a coefficient of 2.97. This means that for a 10% increase in emotional words (an increase of 0.1 in the ratio), there is an 29.7% increase in retweets.

#### Figure 4: Compared: all dictionaries
```{r combined plot, echo = F}
ggplot(gret_sent) +
  geom_smooth(aes(bingRatio, log(retweet_count), color = "red"), method = "lm", se=FALSE) +
  geom_smooth(aes(loughRatio, log(retweet_count), color = "green"), method = "lm", se=FALSE) +
  geom_smooth(aes(nrcRatio, log(retweet_count), color = "blue"), method = "lm", se=FALSE) +
  xlab("Ratio Measure (Dictionary Words: Total Tweet Words)") +
  ylab("Retweet Count (Log)") +
  ggtitle("How Emotionally Laden Words Affect Retweet Count") + 
  scale_color_discrete(name = "Dictionary", labels = c("Bing", "Lough", "Nrc"))
  
```

Across all three dictionaries, we saw a statistically significant positive relationship between the ratio of emotional words to total words in the tweet and the number of retweets the tweet received. While the coefficients were different, we still have statistically significant evidence that an increase in the ratio of emotionally laden words is related to an increase in a tweet’s retweet count. 


## Conclusion

After our analysis, we concluded that Greta Thunberg’s tweets that have a higher ratio of emotionally-laden words to total words will receive greater interaction, measured by the number of retweets. Our main evidence to support this conclusion is that across all three dictionaries there is a statistically significant positive relationship between the word ratio and the logarithm of the retweet count. We can see all three of the positive linear relationships in Figure 4. In tables 1, 2, and 3, we can see that the positive coefficients are all statistically significant.

One limitation of our analysis is that we could not pull all of the tweets from Thunberg’s Twitter account. Since we had limited access to the Twitter API and not research-level access, we could only use around 3,000 of her most recent tweets, which is only a fraction of her account. 

Another limitation is that we cannot be certain that the sentiment dictionaries include all possible emotional words. This would mean that some tweets with more emotional words may have been given lower ratios in our analysis. 

Another issue that remains uncertain is that we cannot differentiate between retweets that come from real humans and retweets that come from bots. We cannot draw a conclusion that people retweet more emotional tweets more vs. if bots amplify emotional tweets more. This would be interesting to explore in a study with more resources or more time because what tweets bots amplify could have strong real-world impacts.

Another issue that remains uncertain is if the conclusions we drew here apply to all tweets about climate change, which was our original research question. Greta Thunberg’s account was our chosen sample since she is an important climate change spokesperson, but her account does not comprise all tweets about climate change. With more time and more access to the Twitter API, we would be able to investigate the research question further and see if our results are similar across other accounts, hashtags, or topics.

